---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# discretewq

<!-- badges: start -->
[![R build status](https://github.com/sbashevkin/discretewq/workflows/R-CMD-check/badge.svg)](https://github.com/sbashevkin/discretewq/actions)
[![Codecov test coverage](https://codecov.io/gh/sbashevkin/discretewq/branch/main/graph/badge.svg)](https://codecov.io/gh/sbashevkin/discretewq?branch=main)
[![DOI](https://zenodo.org/badge/309747392.svg)](https://zenodo.org/badge/latestdoi/309747392)
[![Data DOI](https://img.shields.io/badge/Data%20publication%20DOI-10.6073/pasta/567ca1dce56cc819b1819117538bd718-blue.svg)](https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=731)
<!-- badges: end -->


The goal of discretewq is to integrate discrete water quality data from the San Francisco Estuary. 

## Installation

You can install the latest version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sbashevkin/discretewq")
```

## Usage

To obtain the full integrated water quality dataset

```{r}
require(discretewq)
Data <- wq(Sources = c("EMP", "STN", "FMWT", "EDSM", "DJFMP",
                       "SDO", "SKT", "SLS", "20mm", "Suisun", 
                       "Baystudy", "USBR", "USGS_SFBS", "YBFMP", "USGS_CAWSC"))

str(Data)
```


## Data publication

The dataset is also [published on the Environmental Data Initiative](https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=731), where you can find detailed metadata. This static version of the dataset corresponds to version 2.3.2 of the R package ([archived on zenodo](https://zenodo.org/record/6390964)). 


## Data sources

CDFW. 2020a. Bay Study data. [ftp://ftp.wildlife.ca.gov/BayStudy/](ftp://ftp.wildlife.ca.gov/BayStudy/).

CDFW. 2020b. Fall Midwater Trawl data. [ftp://ftp.wildlife.ca.gov/TownetFallMidwaterTrawl/FMWT%20Data/](ftp://ftp.wildlife.ca.gov/TownetFallMidwaterTrawl/FMWT%20Data/).

CDFW. 2020c. Summer Townet data. [ftp://ftp.wildlife.ca.gov/TownetFallMidwaterTrawl/TNS MS Access Data/TNS data/](ftp://ftp.wildlife.ca.gov/TownetFallMidwaterTrawl/TNS MS Access Data/TNS data/).

Cloern, J. E., and T. S. Schraga. 2016. USGS Measurements of Water Quality in San Francisco Bay (CA), 1969-2015 (ver. 3.0 June 2017). U. S. Geological Survey data release. [doi:https://doi.org/10.5066/F7TQ5ZPR](https://doi.org/10.5066/F7TQ5ZPR)

Interagency Ecological Program (IEP), L. Damon, T. Tempel, and A. Chorazyczewski. 2020a. Interagency Ecological Program San Francisco Estuary 20mm Survey 1995 - 2020. Environmental Data Initiative. [doi:10.6073/PASTA/DA7269F6B68975232A2665B211E57229](https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=535&revision=2)

Interagency Ecological Program (IEP), L. Damon, T. Tempel, and A. Chorazyczewski. 2020b. Interagency Ecological Program San Francisco Estuary Spring Kodiak Trawl Survey 2002 - 2020. Environmental Data Initiative. [doi:10.6073/PASTA/2EDAAA415ABE672008E0AF7542AA5D31](https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=527&revision=2)

Interagency Ecological Program (IEP), L. Damon, T. Tempel, and A. Chorazyczewski. 2021c. Interagency Ecological Program San Francisco Estuary Smelt Larva Survey 2009 – 2021. Environmental Data Initiative. [doi:10.6073/PASTA/696749029898FEF9AD268435BEE54D3D](https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=534&revision=3)

Interagency Ecological Program (IEP), S. Lesmeister, and J. Rinde. 2020. Interagency Ecological Program: Discrete dissolved oxygen monitoring in the Stockton Deep Water Ship Channel, collected by the Environmental Monitoring Program, 1997-2018. Environmental Data Initiative. [doi:10.6073/PASTA/3268530C683726CD430C81894FFAD768](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.276.2)

Interagency Ecological Program (IEP), M. Martinez, and S. Perry. 2021. Interagency Ecological Program: Discrete water quality monitoring in the Sacramento-San Joaquin Bay-Delta, collected by the Environmental Monitoring Program, 1975-2020. Environmental Data Initiative. [doi:10.6073/PASTA/31F724011CAE3D51B2C31C6D144B60B0](https://portal.edirepository.org/nis/mapbrowse?scope=edi&identifier=458&revision=4)

Interagency Ecological Program (IEP), R. McKenzie, J. Speegle, A. Nanninga, J. R. Cook, J. Hagen, and B. Mahardja. 2020c. Interagency Ecological Program: Over four decades of juvenile fish monitoring data from the San Francisco Estuary, collected by the Delta Juvenile Fish Monitoring Program, 1976-2019. Environmental Data Initiative. [doi:10.6073/PASTA/41B9EEBED270C0463B41C5795537CA7C](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.244.4)

Interagency Ecological Program (IEP), C. Pien, and N. Kwan. 2022. Interagency Ecological Program: Fish catch and water quality data from the Sacramento River floodplain and tidal slough, collected by the Yolo Bypass Fish Monitoring Program, 1998-2021. ver 3. Environmental Data Initiative. [doi:10.6073/pasta/f5c4362f4a1f370723e2b9113432909f](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.233.3)

Interagency Ecological Program (IEP), C. Pien, J. Adams, and N. Kwan. 2021. Interagency Ecological Program: Zooplankton catch and water quality data from the Sacramento River floodplain and tidal slough, collected by the Yolo Bypass Fish Monitoring Program, 1998-2018. ver 2. Environmental Data Initiative. [doi:10.6073/pasta/baad532af96cba1d58d43b89c08ca081](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.494.2)

O’Rear, T., J. Durand, and P. Moyle. 2020. Suisun Marsh Fish Study. [https://watershed.ucdavis.edu/project/suisun-marsh-fish-study](https://watershed.ucdavis.edu/project/suisun-marsh-fish-study).

Schraga, T. S., E. S. Nejad, C. A. Martin, and J. E. Cloern. 2020. USGS measurements of water quality in San Francisco Bay (CA), beginning in 2016 (ver. 3.0, March 2020). U. S. Geological Survey data release. [doi:https://doi.org/10.5066/F7D21WGF](https://doi.org/10.5066/F7D21WGF)

United States Fish And Wildlife Service, C. Johnston, S. Durkacz, and others. 2020. Interagency Ecological Program and US Fish and Wildlife Service: San Francisco Estuary Enhanced Delta Smelt Monitoring Program data, 2016-2020. Environmental Data Initiative. [doi:10.6073/PASTA/764F27FF6B0A7B11A487A71C90397084](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.415.3)

U.S. Geological Survey. 2022. USGS water data for the Nation: U.S. Geological Survey National Water Information System database, accessed September 1, 2022, at
[doi:10.5066/F7P55KJN](https://doi.org/10.5066/F7P55KJN)

USBR, R. Dahlgren, L. Loken, and E. Van Nieuwenhuyse. 2020. Monthly vertical profiles of water quality in the Sacramento Deep Water Ship Channel 2012-2019.
